Fine-scale Twitter Big Data Reveals Neighborhood Inequalities in Mental Health Effects of High Temperatures

### 2021-04-08 doc/temp-sentiment.tex
more work on introduction paragraphs

%Intro Paragraph
% I think this should be more about climate change and mental health broadly, as well as a bit about how income has been theorized to affect vulnerability, but there is no evidence yet.  This can be shorter as well, as we have a lot of detail in the Background section -Matt
Research has shown strong linkages between ambient environmental conditions and mental health, with extreme temperatures  frequently associated with poorer overall mental health. Individual's mental health is strongly associated with income with people in lower income levels frequently experiencing higher stress levels and less able to find help for coping with these stresses. Other research on how people cope with environmental temperature extremes indicate that wealthier are better able to cope with heatwaves, whether from traveling to cooler climes, accessing air conditioning, or other temperature ameliorating strategies that may require access to financial capital. Yet we know relatively little about the interaction between temperature, mental health and income. Furthermore, previous studies that did attempt to examine relationships between environmental extremes and health sometimes aggregated data at scales too coarse to examine whether income may be an ameliorating factor. This study examines the interactions between these three using relatively fine-scale, publicly available data. Our research expands on existing studies of temperature and sentiment - one measure of mental health - by examining finer-scale data and incorporating income data to examine correlations between temperature extremes and sentiment.

@@

\section{Background}
\subsection{Temperature and Mental Health}
% Heatwaves are known to worsen mental health

@@

\subsection{Mental Health and Income}
% AND mental health is strongly associated with income
% 1）Childhood 2）adulthood 3）gender
% Perceived income level

The relationship between income, socioeconomic status (SES), well-being and mental health is one of the strongest established patterns in psycho-social literature (cite: Easterlin 1973, Holzer et al. 1986, Perry 1996, Ng et al. 2014). Much of the research into these patterns focuses on the inequalities and distribution of the effects on individual well-being. Subsequent work on mental health outcomes, one measure of well-being, have established strong associations between SES, income and mental health with the most and least privileged most commonly associated with the best and worse mental health experiences and outcomes, respectively (cites: Stevenson and Wolfers 2008, Ng et al. 2014).

For example, results from a cross-sectional comparative study of socioeconomic factors and use of mental health services by people living in Ontario, Canada and the United States, indicate clear disparities in use of mental health services between those with the highest income levels and lowest mental health morbidity and those with the lowest income levels and highest mental health morbidity - those with higher incomes tended to have lower mental health morbidity issues while utilizing mental health services more than those with lower incomes and higher mental health issues (cite Katz et al. 1997).

Results from a systematic review of the literature on associations between social inequalities and mental health disorders indicate a clear and prevailing link between  one or more indicators of less social privilege and higher prevalence of mental health disorders (cite Fryers et al. 2003), with low income one of the most consistent markers of and associations with increases in common mental health disorders. Collectively, results from the income and mental health literature indicate that socially disadvantaged populations experience significantly more frequent common mental health disorders. Scholars use mental health disparities to indicate the disproportionate amount of mental health disorders among persons of low SES\citep{RN1292}

The established association between poverty and mental health revealed that that income distribution may have a significant influence upon mental health over and above the effect of poverty \citep{HANANDITA201459}.Poverty has been identified as one of the major risk factors in mental health. For instance, some systematic reviews have shown that socioeconomically disadvantaged children and adolescents were 2-3 times more likely to have mental health problems than others\citep{REISS201324}, while the household increases in financial resources are generally associated with an overall reduction of mental health problems\citep{2015Does}. Others have argued that increases in SES can serve as a buffer against the negative impacts of difficult life experiences on mental health, particularly for individuals already under substantive mental stress (cite Kawachi and Berkman 2001). Thus, SES and income are associated with mental health improvements in SES and income can serve to decrease menta the

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{../res/wbgt-income.png}
  \caption{Change in sentiment across census block income groups as the Wet Bulb Globe Temperature increases.  For census blocks at the 5th percentile for income, even mild temperatures are associated with decreases in mood, while at higher temperatures, sentiment continues to decline.  For high-income census blocks at the 95th percentile, on the other hand, temperature must be quite high before a decrease in sentiment is observed.}
  \label{fig:wbgt-income}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{../res/wbgt-race.png}
  \caption{Change in sentiment across census block income groups as the Wet Bulb Globe Temperature increases.  For census blocks at the 5th percentile for income, even mild temperatures are associated with decreases in mood, while at higher temperatures, sentiment continues to decline.  For high-income census blocks at the 95th percentile, on the other hand, temperature must be quite high before a decrease in sentiment is observed.}
  \label{fig:wbgt-income}
\end{figure}

>>>>

%Intro Paragraph
% I think this should be more about climate change and mental health broadly, as well as a bit about how income has been theorized to affect vulnerability, but there is no evidence yet.  This can be shorter as well, as we have a lot of detail in the Background section 
% Cite this high-level study: https://www.nature.com/articles/s41558-018-0102-4/
% Also this review article saying more research is needed https://www.sciencedirect.com/science/article/pii/S0033350618302130
% -Matt

Research has shown strong linkages between ambient environmental conditions and mental health, with extreme temperatures  frequently associated with poorer overall mental health. Individual's mental health is strongly associated with income with people in lower income levels frequently experiencing higher stress levels and less able to find help for coping with these stresses. Other research on how people cope with environmental temperature extremes indicate that wealthier are better able to cope with heatwaves, whether from traveling to cooler climes, accessing air conditioning, or other temperature ameliorating strategies that may require access to financial capital. Yet we know relatively little about the interaction between temperature, mental health and income. Furthermore, previous studies that did attempt to examine relationships between environmental extremes and health sometimes aggregated data at scales too coarse to examine whether income may be an ameliorating factor. This study examines the interactions between these three using relatively fine-scale, publicly available data. Our research expands on existing studies of temperature and sentiment - one measure of mental health - by examining finer-scale data and incorporating income data to examine correlations between temperature extremes and sentiment.

@@

\section{Background}
\subsection{Temperature and Mental Health}
Higher temperatures are associated with many indicators of worsened mental health.  Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May}, and many others have demonstrated the same relationship around the world \citep{Qi2014Dec, Page2006Aug, Likhvar2011Jan}.  In addition to suicide, other research has found effects temperature on hospitalization events related to mental health disorders such as bipolar disorder and schizophrenia \citep{Mulllins2019Dec, Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.

Temperature has been hypothesized to impact mental health through a number of pathways.  Work on biological mechanisms emphasize that there may be mental health effects from maintaining stable body temperature in high heat \cite{Lohmus2018Jul}.  Additionally, recent studies have found that nighttime temperatures affects sleep quality \citep{Obradovich2017May, Mullins2019Dec}

% Heatwaves are known to worsen mental health

@@

\subsection{Mental Health and Income}
% This is good, but probably should be shortened.  Maybe one para showing that income matters, one para discussing pathways, also maybe focus/change phrasing to be more on "vulnerability" -Matt

% AND mental health is strongly associated with income
% 1）Childhood 2）adulthood 3）gender
% Perceived income level

The relationship between income, socioeconomic status (SES), well-being and mental health is one of the strongest established patterns in psycho-social literature \citep{Easterlin1974Jan, holzer1986increased, Perry1996Sep} (I cant find Ng et al. 2014). Much of the research into these patterns focuses on the inequalities and distribution of the effects on individual well-being. Subsequent work on mental health outcomes, one measure of well-being, have established strong associations between SES, income and mental health with the most and least privileged most commonly associated with the best and worse mental health experiences and outcomes, respectively \citep{Sevenson2008Aug}(cites: Ng et al. 2014).

For example, results from a cross-sectional comparative study of socioeconomic factors and use of mental health services by people living in Ontario, Canada and the United States, indicate clear disparities in use of mental health services between those with the highest income levels and lowest mental health morbidity and those with the lowest income levels and highest mental health morbidity - those with higher incomes tended to have lower mental health morbidity issues while utilizing mental health services more than those with lower incomes and higher mental health issues \citep{Katz1997}.

Results from a systematic review of the literature on associations between social inequalities and mental health disorders indicate a clear and prevailing link between  one or more indicators of less social privilege and higher prevalence of mental health disorders \citep{Fryers2003}, with low income one of the most consistent markers of and associations with increases in common mental health disorders. Collectively, results from the income and mental health literature indicate that socially disadvantaged populations experience significantly more frequent common mental health disorders. Scholars use mental health disparities to indicate the disproportionate amount of mental health disorders among persons of low SES \citep{RN1292}

The established association between poverty and mental health revealed that that income distribution may have a significant influence upon mental health over and above the effect of poverty \citep{HANANDITA201459}.Poverty has been identified as one of the major risk factors in mental health. For instance, some systematic reviews have shown that socioeconomically disadvantaged children and adolescents were 2-3 times more likely to have mental health problems than others\citep{REISS201324}, while the household increases in financial resources are generally associated with an overall reduction of mental health problems\citep{2015Does}. Others have argued that increases in SES can serve as a buffer against the negative impacts of difficult life experiences on mental health, particularly for individuals already under substantive mental stress \citep{Kawachi2001Sep}. Thus, SES and income are associated with mental health improvements in SES and income can serve to decrease menta the

@@
